---
title: "R Notebook"
output: html_notebook
---
To address the given task, we'll go through it step by step, writing R code for each part. The task involves analyzing BMI data for Dutch boys aged 10 to 11 years from the `dbbmi` dataset in the `gamlss.data` package, fitting parametric distributions, and selecting an appropriate one based on the analysis.

### 1. Plotting the Data

First, let's load the data, subset it for the specific age group, and plot the histogram to find a suitable value for `nbins`.

```{r}
library(gamlss.data)
library(MASS)
library(gamlss)
library(ggplot2)

# Load the dataset
data(dbbmi)

# Subset for ages 10 to 11
old <- 10
da <- with(dbbmi, subset(dbbmi, age > old & age < old + 1))
bmi10 <- da$bmi

# Plot the histogram
truehist(bmi10, nbins=30) # Adjust nbins as needed to make the histogram look good
```

Experiment with the `nbins` parameter to find a visually appealing and informative histogram. The goal is to have enough bins to clearly see the distribution's shape without making it too noisy.

### 2. Fitting Parametric Distributions

Next, we'll fit several parametric distributions to the data. Common distributions for BMI data include the Normal, Log-Normal, and Gamma distributions, among others. The `gamlss` package provides functions to fit a wide range of distributions.

```{r}
# Fit different distributions
fit_norm <- gamlss(bmi10 ~ 1, family=NO)
fit_lognorm <- gamlss(bmi10 ~ 1, family=LOGNO)
fit_gamma <- gamlss(bmi10 ~ 1, family=GA)

# Compare models
models <- list(fit_norm, fit_lognorm, fit_gamma)
aic_values <- sapply(models, AIC)
print(aic_values)

# Select the model with the lowest AIC
selected_model <- models[[which.min(aic_values)]]
```

The choice of distribution can be justified by comparing the Akaike Information Criterion (AIC) values of the fitted models—the model with the lowest AIC is typically preferred as it suggests a good fit with relatively lower complexity.

### 3. Output Parameter Estimates and Interpretation

Finally, for the chosen model, we can output the parameter estimates and interpret them according to the distribution's characteristics.

```r
# Output parameter estimates for the chosen model
summary(selected_model)
```

Interpretation of the parameters will depend on the selected distribution. For example:
- For a Normal distribution (`NO`), the parameters are the mean (`mu`) and standard deviation (`sigma`), representing the location and scale of the distribution.
- For a Log-Normal distribution (`LOGNO`), `mu` and `sigma` represent the mean and standard deviation of the variable's logarithm, indicating the distribution's central tendency and spread on a log scale.
- For a Gamma distribution (`GA`), the parameters might include a shape and a scale parameter, reflecting the distribution's skewness and scale.

Refer to the GAMLSS book or documentation for specific interpretations of the parameters of your chosen distribution. The interpretation will help in understanding the characteristics of BMI distribution among Dutch boys aged 10 to 11, such as its central tendency, variability, and potential skewness.

## Model Selection Week 8

Broad outline of approach for Q3

### Determining an Appropriate Distribution

1. Exploratory Data Analysis (EDA): Start with EDA to understand the distribution of your response variable and the relationships between explanatory variables and the response. This can include plotting histograms, box plots, and scatter plots.

2. Distribution Selection: Based on the EDA, you can hypothesize which distributions might fit your response variable. GAMLSS supports a wide range of distributions, so consider whether your data suggests normality, skewness, kurtosis, or zero-inflation which could guide your choice towards distributions like Gaussian, Binomial, Poisson, Negative Binomial, etc.

3. Fit Multiple Models: Initially, fit models with different distributions to your data. For instance, if you're unsure whether your data are Poisson or Negative Binomial distributed, fit both models.

4. Distribution Comparison: Use criteria like the Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), or Generalized Akaike Information Criterion (GAIC) to compare models with different distributions. Lower values typically indicate a better fit to the data.

### Selecting Relevant Explanatory Variables

1. Univariate Analysis: Start by fitting simple models that include the response variable and one explanatory variable at a time. This can help identify variables that have no significant relationship with the response.

2. Multivariate Analysis: Include all significant variables in a multivariate GAMLSS model. Be mindful of multicollinearity among explanatory variables, as it can affect model estimates.

3. Variable Selection Techniques: Utilize backward elimination, forward selection, or both in a stepwise fashion to select relevant variables. Functions like stepGAIC can automate this process based on information criteria.

### Model Diagnostics

1. Residual Analysis: Check the residuals of your fitted model for any patterns or systematic deviations from assumptions. Plots like residual vs. fitted values or Q-Q plots are useful.

2. Check for Overdispersion: Particularly for count data, ensure that the model accounts for overdispersion if present. This might influence your choice of distribution.

3. Influence Measures: Assess the influence of individual data points on your model using diagnostic measures like Cook’s distance.

4. Model Fit Statistics: Beyond AIC or BIC, consider using goodness-of-fit tests or R-squared analogs for GAMLSS models to evaluate how well your model explains the data.

### Putting It All Together

1. Iteration: Model building is an iterative process. Based on diagnostics and fit statistics, you may need to revisit your choice of distribution, the set of explanatory variables, or even transform variables to meet model assumptions.

2. Validation: Use cross-validation or split your dataset into training and test sets to validate your model's predictive performance on unseen data.


### Sample Code:

```{r}
# Load the necessary library
library(gamlss)

# Assuming my_data is already loaded into your R environment
# If not, you would load it using something like:
# my_data <- read.csv("path/to/your/my_data.csv")

# Initial exploration to determine a potential distribution family
# This step might involve plotting and basic summaries
# For demonstration, we'll proceed directly to fitting models

# Fit models with different distributions to find the best fit
# Replace GA, NO, and BI with the distributions you're interested in
# Example: GA (Gamma), NO (Normal), BI (Binomial)
models <- list(
  GA = gamlss(response ~ p_1 + p_2 + p_3, family = GA, data = my_data),
  NO = gamlss(response ~ p_1 + p_2 + p_3, family = NO, data = my_data),
  BI = gamlss(response ~ p_1 + p_2 + p_3, family = BI, data = my_data)
)

# Compare models using AIC
sapply(models, AIC)

# Proceed with the model having the lowest AIC value
# For demonstration, let's assume the Gamma distribution is chosen

# Model fitting with the selected distribution
final_model <- gamlss(response ~ p_1 + p_2 + p_3, family = GA, data = my_data)

# Variable selection using stepGAIC
final_model_step <- stepGAIC(final_model, method="backward")

# Model diagnostics
# Residual plots
plot(residuals(final_model_step), type="p") # Basic residual plot

# QQ plot for residuals
qqnorm(residuals(final_model_step))
qqline(residuals(final_model_step))

# Checking for influential observations
plot(influence.measures(final_model_step)$infmat)

# Adjust the model as needed based on diagnostics
```