---
title: "MA7007 - Statistical Modelling and Forecasting Case Study Report 2023-2024"
output:
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# suppress the warnings by setting warn=-1 
options(warn=-1) 
```

First, let's load the data, subset it for the specific age group, and plot the histogram to find a suitable value for `nbins`.

```{r loading_data}
# Load the packages
library(ggplot2)
library(gamlss)
library(gamlss.ggplots)
library(gamlss.add)
library(gamlss.data)
library(MASS)

# Load the dataset
data(dbbmi)
summary(dbbmi)
```

```{r select_age_group}
# Subset for ages 15 to 16
old <- 15
dbbmi_15 <- with(dbbmi, subset(dbbmi, age > old & age < old + 1))
bmi15 <- dbbmi_15$bmi
summary(bmi15)
```

```{r histograms}
# Plot the histogram; adjust nbins as needed to make the histogram look good
binwith = 0.5
nbins = trunc(max(bmi15) - min(bmi15)) / binwith

truehist(bmi15, nbins=nbins)

density(bmi15, cut = 0)

gamlss.ggplots:::y_hist(dbbmi_15$bmi,
                        from=floor(min(bmi15)), 
                        to=ceiling(max(bmi15)),
                        binwidth=binwith,
                        title="Histogram of BMI for 15 year olds")
```
Next, we'll fit several parametric distributions to the data. Common distributions for BMI data include the Normal, Log-Normal, and Gamma distributions, among others. 
The `gamlss` package provides functions to fit a wide range of distributions.

```{r fitDist}
# Sense check our thought on skew and kurtosis by plotting against the Normal distribution
plot(histDist(bmi, "NO", density=TRUE, line.col=c(1,1), line.ty=c(1,2), nbins=nbins, data=dbbmi_15))
```

The choice of distribution can be justified by comparing the Akaike Information Criterion (AIC) values of the fitted models—the model with the lowest AIC is typically preferred as it suggests a good fit with relatively lower complexity.

```{r chooseDist}
# No explanitory variable so we can use fitDist to get the best dist for continuous distributions
f1 <- fitDist(bmi, type='realAll', data=dbbmi_15, k=2)
f1$fits[1:8]

# Validate fitDist by looking at the AIC for different GAIC penalties
m0 <- gamlss(bmi ~ 1, family=NO, data=dbbmi_15)
c1 <- chooseDist(m0, type='realAll', data=dbbmi_15, parallel="snow", ncpus=4)
c1

# Here is out selected fitted distribution 
fit_exgaus <- update(m0, family="exGAUS", k=2)
```

The Best fit is exGAUS

```{r}
# Plot out the distribution with the lowest GAIC
m_exgaus <- histDist(bmi, "exGAUS", density=TRUE, line.col=c(1,1), line.ty=c(1,2), nbins=nbins, data=dbbmi_15)
plot(fit_exgaus)
wp(fit_exgaus)
```

Finally, for the chosen model, we can output the parameter estimates and interpret them according to the distribution's characteristics.


```{r}
# Value of fitted parameters
fitted(f1, "mu")[1]
fitted(f1, "sigma")[1]
fitted(f1, "nu")[1]
```

```{r}
# Output parameter estimates for the chosen model
summary(fit_exgaus)
```

Interpretation of the parameters will depend on the selected distribution.