---
title: "MA7007 - Statistical Modelling and Forecasting Case Study Report 2023-2024"
output:
  pdf_document:
    keep_tex: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# suppress the warnings by setting warn=-1 
options(warn=-1)
```

```{r}
# Load the packages
library(ggplot2)
library(gamlss)
library(gamlss.ggplots)
library(gamlss.add)
library(gamlss.data)
```

Read the data file.

```{r}
# Load the data
data("grip", package = "gamlss.data")
```

Set unique seed number, and select 1000 random rows.

```{r}
set.seed(567) 

# Sample 1000 observations
index <- sample(nrow(grip), 1000)
grip_sample <- grip[index, ]
dim(grip_sample)
```

Plot grip against age. 
    
```{r}
# Plot grip against age
plot(grip$age, grip$grip,
     xlab = "Age",
     ylab = "Grip Strength",
     main = "Grip Strength vs Age",
     col = "blue",
     pch = 19)

# Add a smooth line to highlight the trend
lines(smooth.spline(grip$age, grip$grip), col = "red")

# Plot sample grip against age
plot(grip_sample$age, grip_sample$grip,
     xlab = "Age",
     ylab = "Grip Strength",
     main = "Grip Strength vs Age",
     col = "blue",
     pch = 19)

# Add a smooth line to highlight the trend
lines(smooth.spline(grip_sample$age, grip_sample$grip), col = "red")
```

Use the LMS method to fit the data.

```{r}
# Fit the model
gbccg <- gamlss(grip ~ pb(age),
                sigma.fo = ~ pb(age),
                nu.fo = ~ pb(age),
                data = grip_sample,
                family = BCCG)

# Extract Effective Degrees of Freedom
edf_details <- edfAll(gbccg)

# Print the EDF for each parameter
print(edf_details)
```

Use the fitted values from the LMS model as starting values for fitting the BCT and the BCPE distributions to the data

```{r}
# Fit the BCT model using gbccg as starting values
gbct <- gamlss(grip ~ pb(age),
               sigma.fo = ~ pb(age),
               nu.fo = ~ pb(age),
               tau.fo = ~ pb(age),
               data = grip_sample,
               family = BCT,
               start.from = gbccg)

# Fit the BCPE model using gbccg as starting values
gbcpe <- gamlss(grip ~ pb(age),
                sigma.fo = ~ pb(age),
                nu.fo = ~ pb(age),
                tau.fo = ~ pb(age),
                data = grip_sample,
                family = BCPE,
                start.from = gbccg)
```

Effective degrees of freedom fitted for the parameters.

```{r}
# The EDF indicates the complexity of the model related to each parameter.

# Extract EDF for BCT model
edf_bct <- edfAll(gbct)

# Extract EDF for BCPE model
edf_bcpe <- edfAll(gbcpe)

# Print the EDF for each model
print("BCT EDF")
print(edf_bct)
print("BCPE EDF")
print(edf_bcpe)
```

```{r}
edfAll(gbcpe)
```

GAIC to compare the three models.

```{r}
# Calculate GAIC for each model
gaic_bccg <- GAIC(gbccg)
gaic_gbct <- GAIC(gbct)
gaic_gbcpe <- GAIC(gbcpe)

# Print the GAIC values
print(paste("GAIC for BCCG:", gaic_bccg))
print(paste("GAIC for BCT:", gaic_gbct))
print(paste("GAIC for BCPE:", gaic_gbcpe))
```

Plot the fitted parameters for the fitted models in (d) and (e) using for example

```{r}
# Assuming 'grip_sample' is your dataset and 'gbccg' and 'gbct' are fitted models
fittedPlot(gbccg, gbct, x=grip_sample$age)

# Calculate fitted values or centiles for each model across a range of ages
age_seq <- seq(min(grip_sample$age), max(grip_sample$age), length.out = 100)

# For BCCG Model
fitted_bccg <- predict(gbccg, newdata=data.frame(age=age_seq), type="response")

# For BCT Model
fitted_gbct <- predict(gbct, newdata=data.frame(age=age_seq), type="response")

# Plotting
plot(age_seq, fitted_bccg, type='l', col='blue', ylim=range(c(fitted_bccg, fitted_gbct)),
     xlab='Age', ylab='Fitted Grip Strength', main='Fitted Models Comparison')
lines(age_seq, fitted_gbct, col='red')

# Adding a legend
legend("topright", legend=c("BCCG", "BCT"), col=c("blue", "red"), lty=1, cex=0.8)
```

Centile plot for the fitted models.

```{r}
centilesTwo(gbccg, 10:16, seq(15, 40, 5), age,  grip, cent=0.05, dist=.1, xlab="age", ylab='grip')
centiles(gbccg, xvar=grip_sample$age, cent=c(0.1, 0.4, 2,10, 25, 50,75,90,98,99.6, 99.9), ylab="grip", xlab="age", legend=FALSE)
```

```{r}
# Plot centiles for BCCG model
plot(grip_sample$age, grip_sample$grip, col="gray90", main="Centile Comparison", xlab="Age", ylab="Grip Strength")

centiles(gbccg, xvar = grip_sample$age, col = "blue", lty = 1) #, add = TRUE
legend("topright", legend=c("BCCG"), col=c("blue"), lty=1, cex=0.8)

# Add centiles for BCT model to the existing plot
centiles(gbct, xvar = grip_sample$age, col = "blue", lty = 2)
# Update the legend to include BCT
legend("topright", legend=c("BCT"), col=c("blue"), lty=1:2, cex=0.8)
```

Residuals from the fitted models:

```{r}
# For BCCG Model
plot(residuals(gbccg), main="Residuals for BCCG Model")

# For BCT Model
plot(residuals(gbct), main="Residuals for BCT Model")

# For BCCG Model
wp(gbccg, main="Worm Plot for BCCG Model")

# For BCT Model
wp(gbct, main="Worm Plot for BCT Model")

# For BCCG Model
qstats_bccg <- Q.stats(gbccg)
print(qstats_bccg)

# For BCT Model
qstats_bct <- Q.stats(gbct)
print(qstats_bct)
```
