---
title: "MA7007 Week8 Model Selection"
output: html_notebook
---

```{r}
#===========================
# The US Air Pollution Data
#===========================
library(gamlss)
data(usair)
```


```{r}
mod1 <- gamlss(y ~ ., data = usair, family = GA)
```

The drop1() and add1() functions are used to assess the significance of terms in the model by dropping or adding them.

```{r}
# drop1()

# drop1(mod1)

drop1(mod1, parallel="snow", ncpuss=4)

# add1()

# add1(mod1, scope = ~(x1 + x2 + x3 + x4 + x5 + x6)^2)

add1(mod1, scope = ~(x1 + x2 + x3 + x4 + x5 + x6)^2, parallel="snow", ncpuss=4)
```

The stepGAIC() function is used to perform stepwise model selection using the Generalized Akaike Information Criterion (GAIC). Different options are explored, such as parallelization, setting the k parameter, and specifying the scope of terms to be considered.

```{r}
# stepGAIC()

mod2 <- stepGAIC(mod1, parallel="snow", ncpuss=4)

mod21 <- stepGAIC(mod1, k=length(usair$y))

mod3 <- stepGAIC(mod1, scope=list(lower=~1,
                                  upper=~(x1+x2+x3+x4+x5+x6)^2))
mod2$anova

mod21$anova

mod4 <- stepGAIC(mod1, parameter="sigma",
                 scope=~x1+x2+x3+x4+x5+x6)
```

The GAIC() function is used to compare the fitted models based on their GAIC values.

```{r}
GAIC(mod1, mod2, mod21, mod3, mod4) # model selection
```

The plot() and wp() functions are used for diagnostic plots and worm plots, respectively, for the selected models.

```{r}
## plot() and wp(): # diagnostics

plot(mod3)

wp(mod3)
```


The stepGAICAll.A() and stepGAICAll.B() functions are used for automated model selection, considering linear terms and smoothers.

```{r}
# stepGAICAll.A()

m1 <- gamlss(y~1, data=usair, famly=GA,
             trace=FALSE) # null model

m2 <- stepGAICAll.A(m1, scope=list(lower=~1,
                                   upper=~x1+x2+x3+x4+x5+x6))# linear terms

m3 <- stepGAICAll.A(m1, scope=list(lower=~1,
                                   upper=~pb(x1)+pb(x2)+pb(x3)+pb(x4)+pb(x5)
                                   +pb(x6)), k=log(41)) # smoothers

# stepGAICAll.B()
m4<- stepGAICAll.B(m1, scope=list(lower=~1,
                                  upper=~x1+x2+x3+x4+x5+x6),
                   k=log(41))
```



```{r}
AIC(m1, m2, m3, m4)

plot(m4)
wp(m4)
```


## AIDS

```{r}
#======================================================
# The AIDS data cases in the U.K. from January 1983
# to March 1994
# y : number of quarterly aids cases in England & Wales
# x : time in quarters from January 1983
# qrt : a factor for the quarterly seasonal effect
#======================================================

library(gamlss)

data(aids)
```

The find.hyper() function is used to estimate the hyperparameters of the model using a grid search and the Global GAIC criterion.

```{r}
# Global GAIC: find.hyper()

mod1 <- quote(gamlss(y ~ cs(x, df = p[1]) + qrt,
                     family = NBI, data = aids, 
                     trace = FALSE))

op <- find.hyper(model = mod1, par = c(3),
                 lower = c(1), steps = c(0.1), pen = 2.8)
```

Local likelihood estimation methods are explored, including local Maximum Likelihood (ML), local Generalized Cross-Validation (GCV), and local GAIC, using the pb() function for penalized B-splines.

```{r}
# Local ML

data(aids)
m1 <- gamlss(y~pb(x)+qrt, data=aids, family=NBI)

edf(m1)

# Local GCV

m2 <- gamlss(y~pb(x, method="GCV")+qrt, data=aids, family=NBI)

edf(m2)

# Local GAIC

m3 <- gamlss(y~pb(x, method="GAIC", k=2)+qrt, data=aids, family=NBI)
```

The edf() function is used to obtain the effective degrees of freedom for the fitted models, and the GAIC() function is used again for model comparison.

```{r}
edf(m3)

GAIC(m1, m2, m3)
```

Diagnostic plots are generated for the selected model

```{r}
## plot () and wp()
plot(m3)

wp(m3)
```
